### 2023-06-30 includes/aux/final_checks.tex
File structure changes

%%%%%%%%%%%%%%%%%%%%%%%%%%
% Final To-Dos and Checks:
%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{sambox}{Final Double Checks:}
\begin{itemize}
    \item
    No comments left (\texttt{ctr-f} comment commands);

    \item
    No ``\textbackslash iffalse ... \textbackslash fi'' statements or ``...''s.

    \item 
    No \texttt{?}s left in citations, references to Sections, etc. (\texttt{ctr-f} `?' in .pdf);

    \item
    No files in temp folders;

    \item
    Consistency for
    \begin{itemize}
	\item
	high energy (low energy) vs. high-energy (low-energy)

	\item
	``nonperturbative'' to ``non-perturbative''

	\item 
	``subjet'' to ``sub-jet''.

	\item
	\(\beta\) function \(\to\) beta function (so people can find with ctr-f)

	\item
	``plus-function'' to ``plus-distribution''

	\item \(d\) to \(\dd\) in integrals
    \end{itemize}
\end{itemize}
\end{sambox}