### 2023-07-03 chapters/particles.tex
Fixed some small bugs, commented out abstract.

\chapter{Particles Inside Particles:\\Jets and Jet Substructure}

% ==============================================
\section{The physical picture}
% ==============================================

% -----------------------------------
% Figure Comment:
% -----------------------------------
\begin{figure}[t!]
    % Figure graphics
    \centering
    \includegraphics[width=\textwidth]{figures/tempfig}

    % Caption
    \caption{
        Caption text
    }

    % Figure Label
    \label{fig:label}
\end{figure}

% -----------------------------------
% ---------------------------------------
\subsection{Splitting Functions}
% ---------------------------------------

% ---------------------------------------
\subsection{Renormalization}
% ---------------------------------------

% ==============================================
\section{}
% ==============================================