### 2023-07-03 chapters/qcd.tex

\chapter{Quantum Chromodynamics:\\A Picture Book}