### 2023-07-03 chapters/jet_substructure.tex
Adjusted chapters, new jet substructure chapter.

\chapter{Removing the Rubble:\\Jet Substructure, Grooming, Pileup Mitigation, and Cleanup Fish}

% ==============================================
\section{The physical picture}
% ==============================================

% -----------------------------------
% Figure Comment:
% -----------------------------------
\begin{figure}[t!]
    % Figure graphics
    \centering
    \includegraphics[width=\textwidth]{figures/tempfig}

    % Caption
    \caption{
        Caption text
    }

    % Figure Label
    \label{fig:label}
\end{figure}
% -----------------------------------

% ==============================================
\section{Jet Substructure}
% ==============================================

% ==============================================
\section{Jets, Data, and Noise}
% ==============================================

% ---------------------------------------
\subsection{Jet Grooming and Pileup Mitigation}
% ---------------------------------------

% ==============================================
\section{Continuity as a bridge between grooming and pileup mitigation}
% ==============================================

% ==============================================
\section{Pileup and Infrared Annihilation (PIRANHA): A paradigm for continuous jet grooming}
% ==============================================

% ==============================================
\section{Alternatives to Grooming}
% ==============================================

% ==============================================
\section{Connecting back out}
% ==============================================