### 2023-07-03 chapters/particles.tex

\chapter{Particles Inside Particles:\\Jets and Jet Substructure}

% ==============================================
\section{The physical picture}
% ==============================================

% -----------------------------------
% Figure Comment:
% -----------------------------------
\begin{figure}[t!]
    % Figure graphics
    \centering
    \includegraphics[width=\textwidth]{figures/tempfig}

    % Caption
    \caption{
        Caption text
    }

    % Figure Label
    \label{fig:label}
\end{figure}

% -----------------------------------

% ---------------------------------------
\subsection{Splitting Functions}
% ---------------------------------------

% ---------------------------------------
\subsection{Renormalization}
% ---------------------------------------

% ==============================================
\section{Confinement Project?}
% ==============================================

% ==============================================
\section{Jets and Renormalization}
% ==============================================

% ==============================================
\section{Connecting back out}
% ==============================================

% ==============================================
\section*{Chapter Appendix: More on distributions}
% ==============================================

% ---------------------------------------
\subsection{Definitions}
% ---------------------------------------

% ---------------------------------------
\subsection{Derivations}
% ---------------------------------------

% ---------------------------------------
\subsection{Diagrammatica}
% ---------------------------------------

% ==============================================
\section*{Chapter Appendix: Derivation of the splitting functions}
% ==============================================

% ---------------------------------------
\subsection{Picture and Sketch}
% ---------------------------------------

% ---------------------------------------
\subsection{Proof}
% ---------------------------------------

% ==============================================
\section*{Chapter Appendix: Derivation of the DGLAP equations}
% ==============================================

% ---------------------------------------
\subsection{Picture and Sketch}
% ---------------------------------------

% ---------------------------------------
\subsection{Proof}
% ---------------------------------------