### 2023-12-30 includes/glossary.tex

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Thesis Glossary
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \providecommand{\myglsindextext}[1]{\glsentrydesc{#1} (\glsentrytext{#1})}
% \providecommand{\myglsindex}[1]{\index{\myglsindextext{#1}}}
% \providecommand{\mygls}[1]{\gls{#1}\myglsindex{#1}}
% \providecommand{\myglspl}[1]{\glspl{#1}\myglsindex{#1}}
% \providecommand{\myGls}[1]{\Gls{#1}\myglsindex{#1}}
% \providecommand{\myGlspl}[1]{\Glspl{#1}\myglsindex{#1}}

% =====================================
% Acronyms
% =====================================
% TODO: Figure out how to add descriptions and put into glossary

\newacronym{apollonius}{P-AS}{Apollonius Subtraction}

\newacronym{ca}{C/A}{Cambridge/Aachen}

% CAESAR and description

% DGLAP

\newacronym{eec}{EEC}{Energy-Energy Correlator}

\newacronym{emd}{EMD}{Energy Movers' Distance}

\newacronym{ewoc}{EWOC}{Energy Weighted Observable Correlator}

\newacronym{ivs}{IVS}{Iterated Voronoi Subtraction}

\newacronym{ircsafety}{IRC Safety}{Infra-Red and Collinear Safety}

\newacronym{mmdt}{mMDT}{modified Mass Drop Tagger}

\newacronym{piranha}{PIRANHA}{\textbf{P}ileup and \textbf{I}nfrared \textbf{R}adiation \textbf{A}n\textbf{N}i\textbf{H}il\textbf{A}tion}

\newacronym{qcd}{QCD}{Quantum Chromodynamics}

\newacronym{qft}{QFT}{Quantum Field Theory}

% TODO: add fsoft?
\newacronym{rs}{P-RS}{Recursive Subtraction}

\newacronym{rsf}{P-RSF}{Recursive Subtraction with a Fraction}

\newacronym{sd}{SD}{Soft Drop De-Clustering Algorithm}

\newacronym{scet}{SCET}{Soft-Collinear Effective Theory}

% pT

% C12

% =====================================
% Symbols
% =====================================
\newglossaryentry{as}
{
  name={\ensuremath{a_s}},
  description={
  Another version of the strong coupling constant, \ensuremath{a_s = }\gls{alphas}\ensuremath{/4\pi}, which is often used to remove some common factors of \(4\pi\) arising in QCD calculations
},
}

\newglossaryentry{alphas}
{
  name={\ensuremath{\alpha_s}},
  description={
      The dimensionless strong coupling constant, which is a measure of the strength of the strong force.
      %
      In terms of the parameter \(g_s\) in the QCD Lagrangian, it is \(\alpha_s = \frac{g_s^2}{4\pi}\).
      %
      The effective value of the coupling in an interaction depends on the energy scales in the interaction, as governed by the QCD \gls{betafunction}
},
}

\newglossaryentry{betafunction}
{
    name={\ensuremath{\beta}},
    description={
    }
}

\newglossaryentry{betafunctioncoefficient}
{
    % DEBUG: subscript on beta function coeff?
    name={\ensuremath{\beta_k}},
    description={
    }
}

\newglossaryentry{betasd}
{
    name={\ensuremath{\beta_\text{SD}}},
    description={
        The soft drop angular exponent:
        %
        a free parameter of the \gls{sd} algorithm which which controls its angular dependence.
    }
}

\newglossaryentry{fsoft}
{
    name={\ensuremath{f_{\text{soft}}}},
    description={
        The fraction of the jet energy that is removed by \gls{rsf}.
        %
        % The \gls{rsf} algorithm is described in \Section~\ref{sec:rsf}.
    }
}

% =====================================
% Glossary of Terms
% =====================================

% TODO: Add all glossary terms
\newglossaryentry{additivecontamination}
{
    name=additive contamination,
    description={
    }
}

% additive observables

\newglossaryentry{angularclusteringdiscont}
{
    name=angular clustering discontinuity,
    description={
    }
    % Link to clustering discont??
}

\newglossaryentry{angularcontinuity}
{
    name=angular continuity,
    description={
    }
}

\newglossaryentry{angularity}
{
    name=angularity,
    description={
    }
}

\newglossaryentry{angularordering}
{
    name=angular ordering,
    description={
    }
    % Link to clustering history?
    % Link to Cambridge/Aachen?
}

% anti-kt

\newglossaryentry{binaryjettree}
{
    name=binary jet tree,
    description={
    }
    % Link to clustering history
}

\newglossaryentry{clusteringhistory}
{
    name=clustering history,
    description={
    }
    % Link to de-clustering algorithm
}

\newglossaryentry{clusteringdiscont}
{
    name=clustering discontinuity,
    description={
    }
    % Link to de-clustering algorithm
    % Link to clustering history
    % link to angulary clustering discontinuity
}

\newglossaryentry{collinearsing}
{
    name=collinear singularity,
    description={
    }
}

% TODO: make cs acronym
\newglossaryentry{cs}
{
    name=Constituent Subtraction,
    description={
    }
    % Link to pileup mitigation
}

\newglossaryentry{continuousgrooming}
{
    name=continuous grooming,
    description={
    }
}

% critical emission (link to subsequent and pre-critical)

\newglossaryentry{declusteringalg}
{
    name=de-clustering algorithm,
    description={
    }
}

% Distribution
% disambiguate

\newglossaryentry{eventshape}
{
  name=event shape,
  description={}
}

\newglossaryentry{energyflow}
{
  name=energy flow,
  description={}
}

\newglossaryentry{fastjet}
{
    name=\texttt{FastJet},
    description={
    }
}

\newglossaryentry{fragmentation}
{
  name=fragmentation,
  description={}
}

\newglossaryentry{fragfn}
{
  name=fragmentation function,
  description={}
}

%GECF

%GMBE

% groomed energy fraction

\newglossaryentry{frozencoup}
{
  name=frozen coupling,
  description={}
}

\newglossaryentry{hadronization}
{
  name=hadronization,
  description={}
}

\newglossaryentry{hardcutoffgrooming}
{
  name=hard cutoff grooming,
  description={}
}

% inverse transform method

\newglossaryentry{jet}
{
  name=jet,
  description={}
}

\newglossaryentry{jetcalculus}
{
  name=jet calculus,
  description={}
}

\newglossaryentry{jetclustering}
{
  name=jet clustering,
  description={}
}

\newglossaryentry{jetgrooming}
{
  name=jet grooming,
  description={}
}

\newglossaryentry{jetsubstructure}
{
  name=jet substructure,
  description={}
}

% kt algorithm

% TODO: Make MPI acroynm
\newglossaryentry{mpi}
{
  name=multiple parton interactions,
  description={}
}

% non-global observable, non-global log

% LO, LL, NLO, NLL, etc.
% MLL

% TODO: pdf acronym as well
\newglossaryentry{pdf}
{
  name=parton distribution function,
  description={}
}

\newglossaryentry{partonshower}
{
  name=parton shower,
  description={}
}

\newglossaryentry{pileup}
{
  name=pileup,
  description={}
}

\newglossaryentry{pileupmitigation}
{
  name=pileup mitigation,
  description={}
}

\newglossaryentry{pusubevent}
{
  name=pileup subtracted event,
  description={}
}

\newglossaryentry{plusdistreg}
{
  name=plus-distribution regularization,
  description={}
  % Link to plus-function
}

\newglossaryentry{pythia}
{
  name=\texttt{Pythia},
  description={\sam{version info}}
}

\newglossaryentry{radiator}
{
  name=radiator,
  description={}
}

\newglossaryentry{resummation}
{
  name=resummation,
  description={}
}

\newglossaryentry{softdistortion}
{
  name=soft distortion,
  description={}
}

\newglossaryentry{splittingfn}
{
  name=splitting function,
  description={}
  % Link to DGLAP
}

\newglossaryentry{softcontinuity}
{
  name=soft continuity,
  description={}
}

\newglossaryentry{softsing}
{
  name=soft singularity,
  description={}
}

% sudakov factor/exponent

\newglossaryentry{sumrule}
{
  name=sum rule,
  description={}
}

\newglossaryentry{track}
{
  name=track,
  description={}
}

\newglossaryentry{trackfn}
{
  name=track function,
  description={}
}

% TODO: Make UE acronym
\newglossaryentry{ue}
{
  name=underlying event,
  description={}
}

% UE-corrected distribution
% Link to UE

% veto algorithm
% link to sudakov factor/exponent

% veto region
% link to radiator

\newglossaryentry{zcut}
{
  name=\ensuremath{z_{\text{cut}}},
  description={}
}