### 2025-04-02 figures/diagrams/frules/all_frules.tex

\begin{answer}
\begin{center}
{\normalfont\Large\bfseries\sffamily The Rules for the Gluon Sector of QCD are:}
\end{center}

\begin{subequations}
\input{figures/diagrams/frules/gluon_propagator.tikz}
\input{figures/diagrams/frules/gluon_3pt.tikz}
\input{figures/diagrams/frules/gluon_4pt.tikz}
\end{subequations}
with the counterterm diagrams
\begin{subequations}
% \input{}
\end{subequations}
\end{answer}

\begin{answer}
\begin{center}
{\normalfont\Large\bfseries\sffamily The Rules for the Quark Sector of QCD are:}
\end{center}

\begin{subequations}
\input{figures/diagrams/frules/quark_propagator.tikz}
\input{figures/diagrams/frules/quark_charge.tikz}
\end{subequations}
with the counterterm diagrams
\begin{subequations}
% \input{}
\end{subequations}
Let's also not forget that any time we have a closed fermion loop, the associated diagram gets an additional factor of \(-1\).
\end{answer}

\begin{answer}
\begin{center}
{\normalfont\Large\bfseries\sffamily The Rules for the Ghost Sector of QCD are:}

(in covariant gauges)
\end{center}

\begin{subequations}
\input{figures/diagrams/frules/ghost_propagator.tikz}
\input{figures/diagrams/frules/ghost_charge.tikz}
\end{subequations}
with the counterterm diagrams
\begin{subequations}
% \input{}
\end{subequations}
\end{answer}