### 2025-04-02 figures/diagrams/frules/main_frules.tex

\begin{answer}
\begin{center}
{\normalfont\Large\bfseries\sffamily The Rules for the Gluon Sector of QCD are:}
\end{center}

\begin{subequations}
\input{figures/diagrams/frules/gluon_propagator.tikz}

(in a covariant \(R_\xi\) gauge)

\input{figures/diagrams/frules/gluon_3pt.tikz}
\input{figures/diagrams/frules/gluon_4pt.tikz}

External gluons, much like the photons describing the electromagnetic field, are \textit{polarized}.
%
Rather than coming with a propagator factor, as above, external gluons are accompanied by polarization vectors:
\input{figures/diagrams/frules/gluon_external.tikz}
\end{subequations}
\sam{check}
\end{answer}

\begin{answer}
\begin{center}
{\normalfont\Large\bfseries\sffamily The Rules for the Quark Sector of Massless QCD are:}
\end{center}

\begin{subequations}
\input{figures/diagrams/frules/massless_quark_propagator.tikz}
\input{figures/diagrams/frules/quark_charge.tikz}

External quarks have specific spin or \textit{helicity}.
%
External quarks are not associated with the propagator given above, but instead with helicity factors:
\input{figures/diagrams/frules/quark_external.tikz}
\end{subequations}
\sam{check}
\end{answer}