### 2025-04-02 frontmatter/acknowledgements.tex

\section*{Acknowledgements}
% \markboth{Acknowledgements}{Acknowledgements}

%\epigraph{``Understanding is love's other name.''}{Thich Nhat Hanh}
%%
%\epigraph{``We're philosophers. We think, therefore we am.''}{\textit{Small Gods}, Terry Pratchett}
% \epigraph{Films live or die on their casting.}{\href{https://youtu.be/W5bB7-jg_4M?list=PLomTW5B3fAMXJsqE_HDiMpdOdcfAk_Lqc&t=146}{Peter Jackson}}

My first acknowledgements are to my tireless advisor, Jesse Thaler.
%%
%Jesse, you were there with me and for me at MIT from the beginning.
%%
%I remember the email in which you invited me to MIT, our first couple of meetings where we had several wonderful conversations about physics, and your steadfast assurance that I am an important and welcome member of your group.
%
Thank you for taking me under your vast wing.

Rikab Gambhir
%
Sean Benevedes

My second acknowledgments are to my parents, without whom I would surely not be here.

Tengiz Bibilashvili
%
Seth Koren
%
Nathaniel Craig
%
Wouter Waalejwin

Liam Bradley
%
Noah Mihelic
%
Daniel Kreizberg
%
Nate Gardner
%
Kyreija Lamothe
%
Vince Fleming
%
Ryan Croy
%
Rose
%
Erin Neitzel
%
Jason Georgis
%
Johannes Klein
%
Nicholas Maguid

David Rower
(his parents, Nathan Rower)
%
Nick O'Dea
%
James Ehrets
(his parents)
%
David Newsom
%
Jasmin Kwak
%
Ben Siegel
%
Dolev Bluvstein
%
J Rogers

Yurim Lee
%
Kai Goh
%
Eugene Ma
%
Brandon Louie
%
other breakers?
%
Alvin Ong
%
Anoop Praturu
%
Ryan Stuart
%
Marvin Qi

Dimitra Pefkou
%
Jamison Sloan
%
Asmund Folkestad
%
Sarah Geller
%
Yitian Sun
%
Gregory Ridgeway
%
Robert Jones

Neeraj Tata
%
Lamia Ateshian
%
Sarah
%
George
%
Miguel
%
Nathan Everett and Lubah
%
Adam Platek
%
Zhiquan Sun, who taught me that a happy life means eating well.
%
Patrick Oare, Wenzer Qin, Artur Avkadiev
%
Nick Kamp, Caolan John
%
Ouail Kitouni
%
Marianne Moore, Anjie Gao, Manu Srivasta
%
Joshua Lin
%
Atakan Firat, who never backs down from a fight.
%
Lisa Yang, whose constant drive for self-improvement is always inspiring

Jasper Pont, who I still miss.