### 2025-04-02 frontmatter/preface.tex

\section*{Preface}
% \markboth{Preface}{Preface}

\sam{Goal: in a couple of (concerted) weeks, you can master what took me six years (to the extent that I have mastered it)!}

\begin{itemize}
    \item
        Chapters/sections XXX are good for learning

    \item
        My main contributions presented in this thesis are the introduction of \textsc{Piranha} as a paradigm for continuous grooming (\Secs{}{}), the development of new, efficient, and visually intuitive tools for higher-point energy correlators (\Sec{}), and the introduction of the theory of Energy-Weighted Observable Correlations (EWOCs) as a generalization of energy correlators (\Sec{}).
\end{itemize}

\sam{Address Jesse's concerns}

% Include a link to the thesis template

First, and most importantly, please remember that problems are a \textit{shortcut} to understanding, not ``something extra''.
%
At least, good problems are :D
%
I've tried to select and invent some good problems for you (and if you have any comments or qualms with any of them, you should feel free to email me) that I hope are representative of
\begin{itemize}
    \item
        the story or ``flavor'' of the material;

    \item
        the physical concepts, and the intuition I have for the material;

    \item
        the mathematics required to do some realistic (albeit simple) computations.\footnote{Even some numerical calcules! See \Sec{}}
\end{itemize}
%
From what I've seen and read, active learning is the quickest and only way.
%
Don't forget!