### 2025-04-02 frontmatter/submission_information.tex

\begin{center}
    ~
    \vspace{50pt}

    {\LARGE \textbf{Particles Inside Particles:} }

    \vspace{10pt}

    {\LARGE \textbf{The Flow of Energy in Quarks, Gluons, and Jets} }

    \Large

    \vspace{30pt}

    by

    \vspace{25pt}

    Samuel Alipour-fard

    \vspace{40pt}
    % \hrule
    \vspace{40pt}

    Submitted to the Department of Physics in partial fulfillment of the requirements for the degree of

    \vspace{20pt}

    DOCTOR OF PHILOSOPHY

    \vspace{15pt}

    at the

    \vspace{15pt}

    MASSACHUSETTS INSTITUTE OF TECHNOLOGY

    \vspace{20pt}

    May 2025

    \vspace{40pt}
    % \hrule
    \vspace{40pt}

    \copyright{} 2025 Samuel Alipour-fard. All rights reserved.

\end{center}