### 2025-04-09 frontmatter/acknowledgements.tex

\section*{Acknowledgements}
% \markboth{Acknowledgements}{Acknowledgements}

\epigraph{``Understanding is love's other name.''}{Thich Nhat Hanh}
\vspace{-20pt}
\epigraph{``We're philosophers. We think, therefore we am.''}{\textit{Small Gods}, Terry Pratchett}
\vspace{-20pt}
\epigraph{Films live or die on their casting.}{\href{https://youtu.be/W5bB7-jg_4M?list=PLomTW5B3fAMXJsqE_HDiMpdOdcfAk_Lqc&t=146}{Peter Jackson}}

I try as often as I can to remember that we live on a miraculous, miniscule blue dot floating in a vast expanse of emptiness, with scant resources and an incredible story.
%
It is the joy of my life to be part of the humanity sector of team earth.

\vspace{10pt}
\hrule

\begin{center}
Like most of the humanity sector, I am nothing without my mentors.
\end{center}

My first acknowledgements are to my tireless advisor, Jesse Thaler.
%
Thank you for taking me under your vast wing.
%
I have been very lucky to learn from your mastery over every aspect of the scientific process.
%
I have never seen someone do and change so much in the world in such a short time.

My second acknowledgments are to my parents, without whom I would surely not be here.
%
My mother taught me, through her shining example, the values of kindness, openness, acceptance, and sharp awareness and fighting spirit.
%
My father taught me by lack of example the importance of \glslink{jet-grooming}{grooming}, discussed further in \Chap{grooming}.
%
He also gave me the gifts of critical thinking, his jeans, and uncountable priceless lessons, experiences, and memories.

Tengiz Bibilashvili, my academic advisor during my time as an undergraduate at the University of California, Santa Barbara (UCSB), is one of the sharpest, kindest, brightest, and most generous human beings I have ever met.
%
I am grateful every day that I was lucky enough to experience his teaching.

Nathaniel Craig, my research advisor at UCSB, is a powerhouse of intellect, creativity, and pedagogy;
%
I benefited deeply from his dedication to mentorship and patience, and would have been lost on the tip of the iceberg of quantum field theory without his guidance (though, as he would always say with a smile, he was ``just doing his job'').

There is not enough I can say about the mentorship of Seth Koren.
%
Seth is incredibly empathetic and creative, and a brilliant teacher.
%
Seth is, to me, a quintessential human being:
%
a philosopher, clever and kind, who accepts nothing but the truth.

Finally, Wouter Waalejwin, who I met in the fourth year of my PhD, has been an incredibly patient mentor and a valuable resource for me in learning the ins and outs of quantum chromodynamics.
%
His presence is felt in many subtle ways throughout this thesis.
%
I can see his gentle, knowing smile even now.

I would also like to thank Iain Stewart and Christoph Paus.
%
They are members of my thesis committee, and I am grateful for their time and effort on my behalf, but I am most grateful for the environment they created at MIT through their presence.
%
They always had a kind smile to wear, a warm comment to make, or a laugh to share.
%
I knew I could always go to Iain with any question or concern and receive a listening ear and some sound advice.
%
I knew that I could always share a joke and a lighthearted chuckle with Christoph.

On the topic of environment, I would finally like to thank Daniel Harlow.
%
\sam{}

\vspace{12pt}
\hrule
\vspace{12pt}

My sisters, Christina Nazmi-Williams and Mehrnoush Minor (\textit{in alphabetical order}), are my role models and favorite people.
%
It is thanks to their hard work and sacrifices on my behalf that I have been able to live such a clear and irresponsible life.
%
I am so lucky to be in their family.
%
Thomas Minor, my brother-in-law, is another favorite.
%
I aspire to someday be like Tom:
%
brilliant and humble, quietly aware and listening more than speaking, kind and gentle with a troublemaking streak.
%
My nephews and niece -- Ethan and Mason Williams, and Alex and Amelia Minor -- are bundles of joy and diamonds of my life.

My godparents
%
My godsister
%
Last but not least, my godbrother

I also have the best siblings I could ever hope not to be related to.
%
Noah Mihelic (with credit and gratitude to my adopted family: ), who I knew before I had memories, has given me clarity, critical reasoning, incredible companionship, and many magical adventures.
%
David Rower (with credit and gratitude to my adopted family: Rachel Goldberg, Tobin Rower, Nathan Rower, and Lamia Ateshian)
%
J Rogers
%
Neeraj Tata

Sean Benevedes
%
Rikab Gambhir

\vspace{10pt}
\hrule
\vspace{10pt}

The remainder of my acknowledgments are mostly chronological, and no less heartfelt.

Liam Bradley
%
Daniel Kreizberg
%
Nate Gardner
%
Kyreija Lamothe
%
Vince Fleming
%
Ryan Croy
%
Rose
%
Erin Neitzel
%
Jason Georgis
%
Johannes Klein
%
Nicholas Maguid
%
Nathan Everett and Lubah

James Ehrets
(his parents)
%
Nick O'Dea
%
David Newsom
%
Jasmin Kwak
%
Ben Siegel
%
Dolev Bluvstein
%
Anoop Praturu
%
Ryan Stuart
%
Marvin Qi
%
Kai Goh
%
Eugene Ma
%
Brandon Louie
%
other breakers?
%
Alvin Ong
%
Yurim Lee

Zhiquan Sun, who taught me that a happy life means eating well.
%
Shankar Balasubramanian
%
Margarita Davydova
%
Atakan Firat, who never backs down from a fight.
%
Anjie Gao, who taught me (among other things) the wisdom that ``nothing is trivial''.
%
Patrick Oare, Wenzer Qin, Artur Avkadiev
%
Nick Kamp, Caolan John
%
Ouail Kitouni

Dimitra Pefkou
%
Asmund Folkestad
%
Yitian Sun
%
Gregory Ridgeway
%
Jamison Sloan
%
Sarah Geller
%
Robert Jones
%
Anita Budhraja

Lisa Yang, whose constant drive for self-improvement is always inspiring
%
Marianne Moore, Manu Srivasta
%
Joshua Lin
%
Sarah M
%
George
%
Miguel

Jasper Pont, whose free a

\vspace{12pt}
\hrule

\begin{center}
    \textit{Ashes to ashes, dust to dust.}
\end{center}