### 2025-04-10 frontmatter/preface.tex

\section*{Preface}
% \markboth{Preface}{Preface}

\epigraph{Question everything at least once.}{Mansour Alipour-fard}

\epigraph{You earn your right to speculate depending on how much work you do.}{Barton Zweibach}

The nominal goal of this thesis is to present my research from my time in graduate school.

Another goal I have is that, in a couple of (albeit concerted) weeks, you can master what took my six years to learn (to the extent that I have mastered it)!
%
First, and most importantly, please remember that problems are a \textit{shortcut} to understanding, not ``something extra''.
%
At least, good problems are.
%
I've tried to select and invent some good problems for you%
\footnote{
    And if you have any comments or qualms regarding any of them, you should feel free to email me.
}
%
that I hope are representative of
\begin{itemize}
    \item
        the story or ``flavor'' of the material;

    \item
        the physical concepts, and the intuition I have for the material;

    \item
        the mathematics required to do some realistic (albeit simple) computations.
\end{itemize}
%
From what I've seen and read, active learning is the quickest and only way.
%
Don't forget!

Before we begin the thesis, I will add some clarifying comments regarding both goals.
%
First, my main contributions presented in this thesis are
\begin{itemize}
    \item
        The introduction of the \PIRANHA{} paradigm for continuous grooming, with Eric Metodiev, Patrick Komiske, and Jesse Thaler.
        %
        This research was presented in \ul{Pileup and Infrared Radiation Annihilation (PIRANHA): A Paradigm for Continuous Jet Grooming} \cite{Alipour-Fard:2023prj}.
        %
        For \PIRANHA{}, see especially \Secss{piranha}{sd_discont}{grooming-pheno};

    \item
        The development of new, efficient, and visually intuitive higher-point energy correlators -- both \emph{projected} (PENCs) and \emph{resolved} (RENCs) with Ankita Budhraja, Wouter Waalejwin, and Jesse Thaler.
        %
        This research was presented in \ul{New Angles on Energy Correlators} \cite{Alipour-fard:2024szj}.
        %
        For PENCs and RENCs, see \Sec{new-angles};

    \item
        The initiation of the study of general energy-weighted correlations, called Energy Weighted Observable Correlations (EWOCs), that are not restricted to angular correlations (e.g. as in the case of PENCs and RENCs).
        %
        This research was presented in \ul{Energy Correlators Beyond Angles} \cite{Alipour-fard:2025dvp}.
        %
        For EWOCs, see especially \Sec{ewocs}.
\end{itemize}

Second, since I attempt to provide especially clear and step-by-step derivations which highlight at least my own assumptions, several of the derivations presented in this thesis, even of fundamental concepts, are original.
%
Therefore, invoking the uncertainty principle, I advise you to watch closely for their \emph{correctness}.
%
I hope to have given enough detail to make this possible, and I encourage you to email me if I have not.

Finally, in addition to the several texts that appear in my citations below, another under-celebrated text on \glslink{qcd}{quantum chromodynamics} is \underline{Applications of Perturbative QCD}, by Richard D. Field \cite{Field:1989uq}, which pays incredible attention to detail in the presentation of calculations.
%
Of course, it would be a shame to access this book quickly, easily, and for free using \url{libgen.is}, so I will mention that it is also available (at time of writing) at \href{https://www.desy.de/~jung/qcd_and_mc_2009-2010/R.Field-Applications-of-pQCD.pdf}{this desy.de link}.

\begin{center}
Good luck.
\end{center}