### 2025-05-05 frontmatter/submission_information.tex

\begin{center}
    {\LARGE \textbf{Particles Inside Particles:} }

    \vspace{10pt}

    {\LARGE \textbf{The Flow of Energy in Quarks, Gluons, and Jets} }

    \Large

    \vspace{20pt}
    \hruledotted{Submitted}{7.1cm}

    \vspace{-30pt}
    \begin{flushright}
        Samuel Alipour-fard
    \end{flushright}

    \vspace{10pt}

    {\large to the Department of Physics in partial fulfillment of the requirements for the degree}

    \vspace{20pt}

    \textsc{Doctor of Philosophy}  % DOCTOR OF PHILOSOPHY

    \vspace{5pt}

    at the

    \vspace{10pt}

    \textsc{Massachusetts Institute of Technology}  % MASSACHUSETTS INSTITUTE OF TECHNOLOGY

    \vspace{10pt}

    May 2025

    \vspace{10pt}

    {\large \copyright{} 2025 Samuel Alipour-fard. All rights reserved.}
\end{center}